---
title: "Sensitivity Analysis: A4/LEARN Only (No Screen Fails)"
date: "`r Sys.Date()`"
output: 
  html_document:
     prefer-html: true
---

```{r setup, include = FALSE, warning = FALSE, message = FALSE}
library(plyr)
library(nlme)
library(ggplot2)
library(sjPlot)
library(dplyr)
library(gtsummary)
library(flextable)
library(ggrepel)
library(readr)
library(DT)
library(tidyr)

knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
theme_set(theme_bw())
```

```{r}
load(file = "~/Dropbox (Partners HealthCare)/Mabel/Longitudinal A4 & Gene Expression/A4_GeneExp_Project/Data/RNAseqData_Autosomal_XLinked_wDose_121124.RData")

gene_names <- readRDS("~/Dropbox (Partners HealthCare)/Mabel/Longitudinal A4 & Gene Expression/A4_GeneExp_Project/Data/A4_RNAseq_GeneNames_081424.rds")

## Ab Models
load("~/Dropbox (Partners HealthCare)/Mabel/Longitudinal A4 & Gene Expression/A4_GeneExp_Project/Data/Amyloid Models/LME_Ab_GeneE4_NoSF_Models_2025-02-25.RData")
```

```{r}
gene_names <- select(gene_names, gene = ensembl_gene_id, chromosome_name, external_gene_name)
```

The following analyses are re-run versions of our previously found significant results, but this time only including participants from A4 and LEARN (not including participants who screen failed and don't have longitudinal data).

# Significant Results Table

Summary table of all significant results (after fdr correction) from the following longitudinal Amyloid analyses:

1.  Amyloid \~ GeneExp\*E4

```{r}
AbTable_GeneE4_NoSF$predictor <- "Gene*E4"

All_Genes_Sig <- filter(AbTable_GeneE4_NoSF, fdr_p < 0.05)
all_sig_genes <- unique(All_Genes_Sig$gene)
```

```{r}
datatable(All_Genes_Sig,
          filter = "top",
          rownames = FALSE,
          extensions = c('Buttons'),
          options = list(dom = 'Btip',
                         buttons = c('copy', 'csv'),
                         pageLength = 15,
                         columnDefs = list(list(className = 'dt-center', targets = "_all")))) %>%
  formatRound(columns = c("beta","se", "t", "p", "fdr_p"), digits = 4) %>% 
  formatStyle("fdr_p", color = styleInterval(0.05, c("red", "black")))
```

# Amyloid ~ Gene x APOEe4 {.tabset}


```{r}
AbTable_GeneE4_NoSF <- AbTable_GeneE4_NoSF %>%
  left_join(gene_names, by = "gene") %>%
  mutate(Sig = case_when(
    p < 0.05 ~ T,
    TRUE ~ F
  ))
```

There is `r length(subset(AbTable_GeneE4_NoSF, fdr_p < 0.05)$gene)` significant gene that survives FDR correction.

## Volcano Plot

```{r}
sig_bon <- -log10(0.05/length(unique(AbTable_GeneE4_NoSF$gene))) #Bonferroni correction for plotting

ggplot(AbTable_GeneE4_NoSF, aes(x = beta, y = -log10(p), color = Sig)) +
  geom_point() +
  geom_hline(yintercept = -log10(0.05), color = "red", linetype = "dashed", size = 1.2) +
  geom_hline(yintercept = sig_bon, color = "blue", linetype = "dashed", size = 1) +
  theme(legend.position = "none") +
  scale_colour_manual(values = setNames(c('seagreen','black'), c(T,F))) +
  geom_text(x = 0.07, y = 0.9, label = "p = 0.05", check_overlap = TRUE, color = "black") +
  geom_text(x = 0.061, y = (sig_bon-0.5), label = "p = 0.05 \n Bonferroni corrected", check_overlap = TRUE, color = "black") +
  geom_label_repel(aes(label = ifelse(-log10(p) > 5, external_gene_name, "")), 
                   nudge_x = 0, nudge_y = 0.04, min.segment.length = 0) +
  labs(
    title = "Amyloid ~ Gene*APOEe4 Non-Adjusted P-Values"
  )
```

## Summary Table

```{r}
datatable(subset(AbTable_GeneE4_NoSF, Sig == T),
          filter = "top",
          rownames = FALSE,
          extensions = c('Buttons'),
          options = list(dom = 'Btip',
                         buttons = c('copy', 'csv'),
                         pageLength = 15,
                         columnDefs = list(list(className = 'dt-center', targets = "_all")))) %>%
  formatRound(columns = c("beta","se", "t", "p", "fdr_p"), digits = 4) %>%
  formatStyle("fdr_p", color = styleInterval(0.05, c("red", "black")))
```

## Significant Results Plots

### ENSG00000233785

```{r}
rna$APOE <- as.factor(rna$APOE)
rna$SEX <- as.factor(rna$SEX)
rna$Cohort <- as.factor(rna$Cohort)

fit <- lme(suvr_cer ~ ENSG00000233785*APOE*ab_time + bl_ab_age*ab_time + SEX*ab_time + Cohort*ab_time,
           data = subset(rna, SUBSTUDY != "SF"), random =~ ab_time | BID, na.action = "na.omit", 
           control = list(opt = 'optim', msMaxIter = 1000, msMaxEval = 1000))

plot_model(fit, type = "pred", terms = c("ab_time", "ENSG00000233785", "APOE")) +
  theme(strip.background = element_rect(fill = "white")) +
  labs(
    x = "Time (Years)",
    y = "Longitudinal Amyloid",
    title = bquote("Longitudinal Amyloid by SAT1-DT and APOEε4"),
    color = "ENSG00000233785\n(SAT1-DT)"
  )
```

